---
title: "Project 2"
author: "George Bridges and Marcus Lee"
date: "6/27/2021"
output:
  github_document
params:
  day: "Monday"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, comment= FALSE}
# Reading in the libraries/packages
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
library(tidyr)
library(readr)
library(knitr)
library(caret)
```

# `r params$day` Report

## Introduction section ?

This section should briefly describes the data and the variable I will have to work with

Target variable will be the `casual` and/or `registered` variables in some way (perhaps the sum of them, that is up to us)

You should also mention the purpose of the analysis and the methods we will use to model the response.

## Data

After looking at the `day.csv`, the variables not selected for this analysis are listed below:

* `instant`
* `dteday`
* `casual` I will be using the `cnt` variable
* `registered` I will be using the `cnt` variable

```{r, warning = FALSE, message=FALSE}
data_days <- read_csv("day.csv")
```


























## Summarization

## Modeling

## Comparison